---
title: "p8105_hw2_fl2569"
author: "Fang Liu"
date: "10/7/2021"
output: github_document
---

## Problem 1

#### Mr. Trash Wheel Data
Specify the sheet in the Excel file and omit non-data entries using arguments in `read_excel`. 

* Use reasonable variable names 
* Omit rows that do not include dumpster-specific data 
* Round the number of sports balls to the nearest integer 

```{r setup, message = FALSE}
library(tidyverse)
library(readxl)
options(tibble.print_min = 5)
```

```{r data_cleaning, message = FALSE}
trash_wheel_df = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                          sheet = "Mr. Trash Wheel", 
                          range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>%
  mutate(sports_balls = round(sports_balls, digits = 0))
```

#### Precipitation Data
Read and clean the precipitation data for 2018 and 2019:

* Omit rows without precipitation data 
* Add a variable for year

```{r precip, message = FALSE}
#2018 data
precip_2018 =
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
                          sheet = "2018 Precipitation", 
                          skip = 1) %>% 
  janitor::clean_names() %>% 
  mutate(year = 2018) %>% 
  drop_na() 

#2019 data
precip_2019 =
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
                          sheet = "2019 Precipitation", 
                          skip = 1) %>%
  janitor::clean_names() %>% 
  mutate(year = 2019) %>% 
  drop_na()
```

Combine precipitation data sets and convert month (from numeric) to a character variable.
```{r combine}
precip_combined =
  bind_rows(precip_2018, precip_2019) %>% 
  mutate(month = month.name[month])

precip_combined
```

## Problem 2


## Problem 3 